---
title: "Guideline for pottery analysis_example code"
author: "Soren Pedersen"
date: "11/1/2021"
output: html_document
---
---------------
####Prepare####
---------------

### Load Data
```{r}
#Substitute with own data!
load("KEN2020clean.RData")
```
### Load Libraries and install if necessary)
```{r}
# ipak function: install and load multiple R packages.
# https://gist.github.com/stevenworthington/3178163
# check to see if packages are installed. Install them if they are not, then load them into the R session.

ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("ggplot2", "tidyverse", "DescTools", "WVPlots", "Hmisc", "treemapify", "scales", "ggbeeswarm")
ipak(packages)
```

------------------------------
####Descriptive statistics####
------------------------------

We will run all the descriptives. Save these straight into a word file if you are running windows.
```{r}
desc.rims <- Desc(rims)
```

#How many sherds from each site
How many rim shards was registered from each site
```{r}
Desc(rims$Site)
```

Let's do the same with the body shards
```{r}
count_bodyshards <- table(body$Site, body$`No. Of sherds`)
```
...and another method (I think)
```{r}
Desc(body$Site, body$`No. Of sherds`)
```

#Temper
Which type of temper is dominant at the sites
```{r}
Desc(rims$Temper)
```

#Temper by site
```{r}
library(scales)
ggplot(rims, 
       aes(x = factor(Site,
                      levels = c("KEN1", "KEN3","KEN4", "KEN5", 
                                 "KEN6", "KEN7", "KEN8", "KEN9", "KEN10", 
                                 "KEN11", "KEN12")),
           fill = factor(rims$Temper, 
                         levels = c("?", "21", "31", "32", "32c", "33", "34", "35"),
                         labels = c("unknown", "sand only", "grog only", "grog with coarse sand", "grog with coarse sand (big fragments)", "grog with shell/bone", "grog with coarse sand and laterite", "grog with laterite")))) + 
  geom_bar(position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent) +
  scale_fill_brewer(palette = "Set2") +
  labs(y = "Percent", 
       fill = "Temper Category",
       x = "Site",
       title = "Temper Categories by Site") +
  theme_minimal()
```

#Slip
Which type of slip is dominant at our sites and on which location
```{r}
Desc(rims$Slip)
```

#Burnish
Which area is burnished
```{r}
Desc(rims$Burnish)
```

#Slip and burnish
When does slip and burnish occur together and on which parts?
```{r}
#Simple plot
plot(rims$Slip, rims$Burnish)
```
...Don't know whats happening below
```{r}
burnishXslip <- rims %>% 
  group_by(Burnish, Slip) %>% 
  summarise(n=n()) %>% 
  spread(Slip, n) %>% 
  as.data.frame()

burnishXslip$Burnish <- as.character(burnishXslip$Burnish)

burnishXslip <- burnishXslip%>% 
  replace(is.na(.), 0) %>% 
  as.data.frame() %>% 
  column_to_rownames("Burnish")
```

#Decoration
Which type of decoration is dominant at the sites?

Motif 1
```{r}
Desc(rims$Motif1)
```

Motif 2
```{r}
Desc(rims$Motif2)
```

Motif 3
```{r}
Desc(rims$Motif3)
```

Motif 4
```{r}
Desc(rims$Motif4)
```

Continue with these commands if there are more motifs

Lets continue with the body sherds
```{r}
Desc(body$`Motif 1`, valueLabel(body$`No. Of sherds`))
```

#Combine the Motif1-4 categories (frist rims, then body shards)
We start by combining the motif1-4 columns into one column of the rims sheet
```{r}
Motif1_rim <- rims %>% select(Site, Context, Motif1, Loc1) %>% rename (Motif = Motif1, Loc = Loc1)
Motif2_rim <- rims %>% select(Site, Context, Motif2, Loc2) %>% rename (Motif = Motif2, Loc = Loc2)
Motif3_rim <- rims %>% select(Site, Context, Motif3, Loc3) %>% rename (Motif = Motif3, Loc = Loc3)
Motif4_rim <- rims %>% select(Site, Context, Motif4, Loc4) %>% rename (Motif = Motif4, Loc = Loc4)

Motifs_rim <- bind_rows(Motif1_rim, Motif2_rim, Motif3_rim, Motif4_rim) %>% na.omit()

#Then try to plot them

count(Motifs_rim, Motif) %>% plot()

#Total rim count
rimcount <- count(Motifs_rim, Motif)

#Plot the rims

Motifs_rim %>%
  filter(!Motif %in% c("EROD", "PLAIN")) %>% 
  group_by(Motif) %>%
  summarise(n = n())%>%
  mutate(perc = (n/411)*100)%>% 
  filter(perc > 3) %>%
  ggplot(aes(x = Motif, y= perc))+
  geom_bar(stat = "identity")+
  labs(title = "Decorative motifs (rims) on KEN 2020 pottery",
       subtitle = "above 3% prevalence",
       x = "Motif",
       y = "%", 
       caption = "n = 411")+
  theme_minimal()

Desc(Motifs_rim$Motif)
```

Then we combine motif1-4 from the body shard sheet
```{r}
Motif1_body <- body %>% select("Site", "Unit", "Context", "Motif 1", "No. Of sherds") %>% rename (Motif = "Motif 1")
Motif2_body <- body %>% select("Site", "Unit", "Context", "Motif 2", "No. Of sherds") %>% rename (Motif = "Motif 2")
Motif3_body <- body %>% select("Site", "Unit", "Context", "Motif 3", "No. Of sherds") %>% rename (Motif = "Motif 3")
Motif4_body <- body %>% select("Site", "Unit", "Context", "Motif 4", "No. Of sherds") %>% rename (Motif = "Motif 4")

Motifs_body <- bind_rows(Motif1_body, Motif2_body, Motif3_body, Motif4_body) %>% na.omit()

#Then try to plot them

Desc(Motifs_body$Motif)

Motifs_body <- Motifs_body %>% rename(no = `No. Of sherds`)

#Total body count
bodycount <- Motifs_body %>% group_by(Motif) %>% tally(no)

Motifs_body %>%
  filter(!Motif %in% c("EROD", "PLAIN")) %>% 
  group_by(Motif) %>%
  summarise(n = n())%>%
  mutate(perc = (n/611)*100)%>% 
  filter(perc > 3) %>%
  ggplot(aes(x = Motif, y= perc))+
  geom_bar(stat = "identity")+
  labs(title = "Decorative motifs (body) on KEN 2020 pottery",
       subtitle = "above 3% prevalence",
       x = "Motif",
       y = "%", 
       caption = "n = 2599")+
  theme_minimal()

#Total amount of rim+body shards
sum(body$`No. Of sherds`)
2599+411

#Then we combine the rimcount and bodycount

motifs_all <- bind_rows(rimcount, bodycount) %>% group_by(Motif) %>% tally(n)

#And plot the motifs_all

motifs_all %>%
  filter(!Motif %in% c("EROD", "PLAIN")) %>% 
  mutate(perc= (n/3010)*100) %>%
  filter(perc > 2) %>% 
  ggplot(aes(x = Motif, y= perc))+
  geom_bar(stat = "identity")+
  labs(title = "Decorative motifs on KEN 2020 pottery",
       subtitle = "above 2% prevalence",
       x = "Motif",
       y = "%", 
       caption = "n = 3010")+
  theme_minimal()

```

#Diameter
Which opening diameter is preferred?
```{r}
Desc(rims$Diameter)
```

#Average thickness
Which minimum thickness is dominant?
```{r}
Desc(rims$MinTh)
```

And the max neck thickness
```{r}
Desc(rims$MaxTh)
```

#Mean
Calculate the average mean of opening diameter and minimum thickness
```{r}
mean(rims$Diameter, trim = 0, na.rm = TRUE)
```

-------------------------
####Numeric variables####
-------------------------

NOTE: In this part I include the comments because they will be a vital part of interpreting the results. They will be in "" marks.

"We only have three of these, which pertain to the sizes of the vessels: Max Thickness, Min Thickness and Diameter. Since we have individual descriptives, we will now see how they relate to each other and plot this."

# Max to Min Thickness

Gives the relationship of rim to body thickness and tests for correlation.
```{r}
minXmax <- ggplot(data = rims,
                   mapping = aes(x = MinTh,
                                 y = MaxTh))  
minXmax2 <- minXmax +  geom_point(alpha = .7,
                       color = "cornflowerblue") +  #will have to see how to use colours, but this can also be changed via themes
  scale_x_continuous(labels = scales::label_number(suffix = "cm"))+
  scale_y_continuous(breaks = seq(0, 40, 20),
                     label = scales::label_number(suffix = "mm")) + theme_minimal()

minXmaxfull <- minXmax2 +
  stat_summary(fun.data= mean_cl_normal) + 
  geom_smooth(method='lm') +
  labs(title = "Relationship between minimum vessel thickness and maximum rim thickness ",
       subtitle = "all Kéniéroba sites",
       #caption = "2017 Excavations",      #the caption is just an example
       x = "Minimum Thickness",
       y = "Maximum Thickness") 
             
minXmaxfull + theme_minimal()
```

"The regression line is not totally flat, but because of the way the plot is organised, we can't really trust it."

"we can try to do the same with the WVPLots perfect scatterplot, but it might be overkill. It is worth doing this maybe when there is a correlation that we want to show, but not when there isn't."
```{r}
WVPlots::ScatterHist(rims, "MinTh", "MaxTh",
  title= "Relationship of Minimum Vessel Thickness to Maximum Rim Thickness",
  smoothmethod = "lm",
  estimate_sig = TRUE,
  contour = TRUE,
  point_color = "#006d2c", # dark green
  hist_color = "#6baed6", # medium blue
  smoothing_color = "#54278f", # dark purple
  density_color = "#08519c", # darker blue
  contour_color = "#9e9ac8") # lighter purple
```

"This does show us two areas of particular density in the pairings, which has mainly to do with minor peaks in the maxThickness values. The relatively steep regression line also suggests that the values are correlated." 

"In order to not rely on the visual regression line which can be misleading, we will go through the statistical motions here, testing Minimum Thickness against Diameter. The usual method is a Pearson correlation test (Pearson's r), which works on t-tests and thus on the mean values. This goes from 1 (total positive correlation) to -1 (total negative correlation) via 0 (no correlation)."
```{r}
cor.test(rims$MinTh, rims$MaxTh,  method = "pearson", use = "complete.obs")
```
"t is the t-test value, here 8.7, df are degrees of freedom (157), and the p-value is the significance level of the t-test (2.458^{-15}). This last value tells us that if the two samples were not correlated (0), we would get the effect present in our data only in an extremely small number of cases.
Thus we can say that there is a correlation between the two values. cor, the correlation coefficient tells us that they are significantly positively correlated with a value of 0.6, which is a bit more than half-way between not correlated and fully correlated."

# Max Thickness to Diameter

Is the diameter of the opening related to the rim thickness?
```{r}
diamXmax <- ggplot(data = rims,
                   mapping = aes(x = Diameter,
                                 y = MaxTh))  
diamXmax2 <- diamXmax +  geom_point(alpha = .7,
                       color = "cornflowerblue") +
  scale_x_continuous(labels = scales::label_number(suffix = "cm"))+
  scale_y_continuous(breaks = seq(0, 40, 20),
                     label = scales::label_number(suffix = "mm")) + theme_minimal()

diamXmaxfull <- diamXmax2 +
  stat_summary(fun.data= mean_cl_normal) + 
  geom_smooth(method='lm', color="red") +
  labs(title = "Relationship between opening diameter and maximum thickness ",
       subtitle = "all Kéniéroba sites",
       #caption = "2017 Excavations",
       x = "Diameter",
       y = "Maximum Thickness") 
             
diamXmaxfull + theme_minimal()
```

VWplots
```{r}
WVPlots::ScatterHist(rims, "MaxTh", "Diameter",
  title= "Relationship of Maximum Vessel Thickness to Diameter",
  smoothmethod = "lm",
  estimate_sig = TRUE,
  contour = TRUE,
  point_color = "#006d2c", # dark green
  hist_color = "#6baed6", # medium blue
  smoothing_color = "#54278f", # dark purple
  density_color = "#08519c", # darker blue
  contour_color = "#9e9ac8") # lighter purple
```

"This shows us that there is a correlation as the Pearson's R also will show us."

And Pearson's r:
```{r}
cor.test(rims$MaxTh, rims$Diameter,  method = "pearson", use = "complete.obs")
```
"The result is again a singificant correlation, but somewhat weaker than the thickness values to each other."

# Min Thickness to Diameter

Is the diameter of the opening related to the body thickness?
```{r}
diamXmin <- ggplot(data = rims,
                   mapping = aes(x = Diameter,
                                 y = MaxTh))  
diamXmin2 <- diamXmin +  geom_point(alpha = .7,
                       color = "cornflowerblue") +
  scale_x_continuous(labels = scales::label_number(suffix = "cm"))+
  scale_y_continuous(breaks = seq(0, 40, 20),
                     label = scales::label_number(suffix = "mm")) + theme_minimal()

diamXminfull <- diamXmin2 +
  stat_summary(fun.data= mean_cl_normal) + #remove these lines if
  geom_smooth(method='lm') +               #you don't want the regression line
  labs(title = "Relationship between Minimum Vessel Thickness and Opening Diameter",
       subtitle = "all Kéniéroba sites",
       #caption = "2017 Excavations",
       x = "Minimum Thickness",
       y = "Maximum Thickness") 
             
diamXminfull + theme_minimal()
```

VWplots
```{r}
WVPlots::ScatterHist(rims, "MinTh", "Diameter",
  title= "Relationship of Minimum Vessel Thickness to Diameter",
  smoothmethod = "lm",
  estimate_sig = TRUE,
  contour = TRUE,
  point_color = "#006d2c", # dark green
  hist_color = "#6baed6", # medium blue
  smoothing_color = "#54278f", # dark purple
  density_color = "#08519c", # darker blue
  contour_color = "#9e9ac8") # lighter purple
```

Correlation test:
```{r}
cor.test(rims$MinTh, rims$Diameter,  method = "pearson", use = "complete.obs")
```

"These two variables are quite a bit more significantly correlated with a coefficient of 0.7 and p-value of 4207^{-13}." 

"To us this might suggest that **how big an opening diameter might be is related significantly to how thickly the body of the vessel is built**. If you want a bigger opening, your vessel must be thicker." 
-> Idiotic interpretation. Correct: "The larger the vessel is the thicker the walls have to be"

Let's have a look at the 'perfect scatterplot' for this one:
```{r}
WVPlots::ScatterHist(rims, "MinTh", "Diameter",
  title= "Relationship of Minimum Vessel Thickness to Opening Diameter",
  smoothmethod = "lm",
  estimate_sig = TRUE,
  contour = TRUE,
  point_color = "#006d2c", # dark green
  hist_color = "#6baed6", # medium blue
  smoothing_color = "#54278f", # dark purple
  density_color = "#08519c", # darker blue
  contour_color = "#9e9ac8") # lighter purple
```
"We have two peaks again, related to the peaks in the density of the Diameter values."

At the end of this segment, we remove the datasets we have created but no longer need:
```{r}
rm (diamXmax, diamXmax2, diamXmaxfull, diamXmin, diamXmin2, diamXminfull, minXmax, minXmax2, minXmaxfull)
```

## Categorical variables

### Rim Types
To look at a factor that has many levels, like rim types, for a first overview, we can use a treemap: 
```{r}
rimtypes <- count(rims,RimType) # get the data into the right shape with counts of the occurrences.
ggplot(rimtypes, 
       aes(fill = RimType, 
           area = n, 
           label = RimType)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", 
                    place = "centre") +
  labs(title = "KEN2020 Sherds by Rim Type",
       subtitle = "n=411") +
  theme(legend.position = "none")
```

This tells us that ES30 is the most common rim type, followed by S2 and then some T types. We'll take a closer look with desc
```{r}
Desc(rims$RimType)
```

**So ES30 weighs in at 15.3%, S2 at 9.2%. The first 7 rim types, which apart from ES30 all belong to T or S shapes, together account for almost half of the rims (46.3%).**
However, this big picture is still a little confusing. We'll break the rims down into their larger categories, creating a new variable.
```{r}
library(stringi)
library(stringdist)
creating.gentypes <- function(types){
   #the amatch function only works on character
   types <- as.character(types)
   c <- c("S", "ES", "Emg", "Et", "Ef", "Ep", "Esg", "T", "Fl", "F","C", "D", "B", "L", "BT", "I", "Y")
   #amatch will help to correlate the general rimtypes to the more complex ones
    i <- amatch(types, c, maxDist = 8)
    #the code is applied to the data
    rims$RimType.general <- c[i]
    #turned back into a factor 
    types <- as.factor(types)
    rims$RimType.general <- as.factor(rims$RimType.general)
}
rims$RimType.general <- creating.gentypes(rims$RimType)

## we will also save this to the dataset
save(rims, body, file = "KEN2020clean.RData")
```
Then rerun the treemap to see the improvement
```{r}
rimcategories <- count(rims,RimType.general) # get the data into the right shape with counts of the occurrences.
ggplot(rimcategories, 
       aes(fill = RimType.general, 
           area = n, 
           label = RimType.general)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", 
                    place = "centre") +
  labs(title = "KEN2020 Sherds by Rim Type",
       subtitle = "n=411") +
  theme(legend.position = "none")
```

This shows a little more clearly how **everted simple rims and T-rims are the most common**. They seem to make more than half together. Let's check the numbers:
```{r}
Desc(rims$RimType.general)
```

**This shows us that ES has 38.1%, T has 29%, S has 19.1%. ES and T together account for 67.1% of the rims. ES, T and S account for 86.1%. This is a clear tendency towards these three general rim shapes.** But let's not forget that some of the shapes in these categories are amongst the least frequent in the more detailed treemap. Sometimes simplifying data can obscure important findings, too.

#### Rim Types by site
Since we have this dominance of a few rim types, maybe we can find out how they are distributed across the sites that we have surveyed and excavated. Are there differences between the sites?
```{r}
library(scales)
ggplot(rims, 
       aes(x = factor(Site,
                      levels = c("KEN1", "KEN3","KEN4", "KEN5", 
                                 "KEN6", "KEN7", "KEN8", "KEN9", "KEN10", 
                                 "KEN11", "KEN12")),
           fill = factor(RimType.general, 
                         levels = c("ES", "T", "S", "Fl", "B", "Y", "D", "L"),
                         labels = c("simple everted", "thickened", "simple", "flattened", "base",
                                    "Y-shaped", "drainpipe", "lid")))) + 
  geom_bar(position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent) +
  scale_fill_brewer(palette = "Set2") +
  labs(y = "Percent", 
       fill = "Rim Category",
       x = "Site",
       title = "Rim Categories by Site") +
  theme_minimal()
```

Here we see how almost all assemblages are dominated by two rim categories, ES and T. There seems to be quite a difference in the diversity of the assemblages. Only some of them have Y-shaped rims, interestingly it is the three we excavated. KEN4 and KEN9 are the least diverse.

We will have to delve into this in more detail.
```{r}
library(ggpubr)
rimtypeXsite <- xtabs(~ rims$Site + rims$RimType)
attr(rimtypeXsite, "class") <- NULL 
attr(rimtypeXsite, "call") <- NULL 
ggballoonplot(rimtypeXsite, fill = "value")+
  scale_fill_viridis_c(option = "C")
```

too much for a balloon plot.
we can try to visualise this in a correspondence analysis, but it is likely to get messy
```{r}
library(FactoMineR)
library(factoextra)
res.ca <- CA(rimtypeXsite, graph = FALSE)
fviz_ca_biplot(res.ca, repel = TRUE)
```

This is not too bad. KEN5 appears to separate out, while we get KEN1 and KEN10 with similar results. We'll definitely need to stay on this. The next step would be to make a matrix of tables and graphs comparing those sites that are close together on the CA graph. I.e. KEN9 and 7, KEN 1 and 10, KEN 4 and 12, 3 and 11, 8 and 6, and 5.

```{r}
#filter for sites and create new dataset
KEN1_10 <- filter(rims, Site %in% c("KEN1", "KEN10")) %>% count(Site, RimType, sort = TRUE, name = "n")

#then plot
ggplot(KEN1_10, 
       aes(fill = RimType, 
           area = n, 
           label = RimType)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", 
                    place = "centre") +
  labs(title = "Rim Types") +
  theme(legend.position = "none")+
  facet_wrap( ~ Site)
```

We'll keep this going for all sites:
```{r}
rimtype_site <- rims %>%  count(Site, RimType, sort = TRUE, name = "n")

#then plot
ggplot(transform(rimtype_site,
      Site=factor(Site,levels=c("KEN5","KEN8","KEN6", "KEN9", "KEN7", "KEN4", "KEN12", "KEN3", "KEN11", "KEN1", "KEN10"))),#reordered to reflect groups seen in CA 
       aes(fill = RimType, 
           area = n, 
           label = RimType)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", 
                    place = "centre") +
  labs(title = "Rim Types") +
  theme(legend.position = "none")+
  facet_wrap( ~ Site)
```

We'll look at the excavated samples in more detail:
```{r}
KEN8_6 <- filter(rims, Site %in% c("KEN8", "KEN6")) %>% count(Site, RimType, sort = TRUE, name = "n")

#then plot
ggplot(KEN8_6, 
       aes(fill = RimType, 
           area = n, 
           label = RimType)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", 
                    place = "centre") +
  labs(title = "Rim Types") +
  theme(legend.position = "none")+
  facet_wrap( ~ Site)
```

We'll now try to see whether there is a temporal effect here, via seriation (see the seriation.R file):
```{r}
library(tabula)
library(magrittr)
library(tidyverse)
## First, we will look at the sites as sites:
rimtypeXsite <- table(rims$RimType,rims$Site)
rimtypeXsite <- unclass(rimtypeXsite)
rimtypeXsite <- t(rimtypeXsite)

## Coerce dataset to an abundance matrix
#rim_counts <- arkhe::as_count(rimtypeXsite)
## Plot original data matrix
#plot_ford(rim_counts, EPPM = FALSE) +
 # ggplot2::labs(title = "Original dataset") +
  #khroma::scale_fill_bright()
## Get seriation order for columns on EPPM using the reciprocal averaging method
## Expected column order: N, A, C, K, P, L, B, E, I, M, D, G, O, J, F, H
#rim_indices <- seriate_reciprocal(rim_counts, EPPM = TRUE, margin = 2)
## Permute columns
#rim_seriation <- permute(rim_counts, rim_indices)
## Plot new matrix
#plot_ford(rim_seriation, EPPM = TRUE) +
 # ggplot2::labs(title = "Reordered dataset") +
  #khroma::scale_fill_bright()
```

This keeps hanging up with me at the moment. I'll come back to it later...

## Vessel shape
If we want to approach the shape of the vessel, we have three variables to play with: Diameter (numeric), RimAngle (categorical), general Rim Type (categorical). So in order to get an idea of the shape of the vessels, we need to put these three variables together.

First, we create a reduced dataset consisting only of our three dimensions, and remove all rows in which there is an NA. Then we also drop the D rims, because drainpipes are not vessels
```{r}
library(tidyr)
shape <- select(rims, Diameter, RimAngle, RimType.general) %>% drop_na() %>% filter(RimType.general != "D")
```

### Bivariate Analyses
First, we will look at each pair of these values, since there are only three. This will give us a good first impression.
#### RimType and RimAngle
```{r}
ggplot(shape, 
       aes(x = RimAngle, 
           fill = RimType.general)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

ggplot(shape, 
       aes(x = RimType.general, 
           fill = RimAngle)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

```

One very clear result here is that ES rims, who are the most frequent, do not occur on vertical shapes. 

#### Rim Type and Diameter

```{r}
library(ggridges)

ggplot(shape, 
       aes(x = Diameter, 
           y = RimType.general, 
           fill = RimType.general)) +
  geom_density_ridges() + 
  theme_ridges() +
  labs("Diameter by Rim category") +
  theme(legend.position = "none")
```

It seems as though there is a relationship between the rim types and the diameter of the opening. So types thus favour a smaller opening, T types a larger one. Fl has two peaks and ES becomes less prevalent as the diameter gets larger. 

#### Rim Angle and Diameter

```{r}
library(ggridges)

ggplot(shape, 
       aes(x = Diameter, 
           y = RimAngle, 
           fill = RimAngle)) +
  geom_density_ridges() + 
  theme_ridges() +
  labs("Diameter by Rim Angle") +
  theme(legend.position = "none")
```

Here the picture is less clear.

Let's try and express the relationships between Diameter and Type and Angle in a statistical model, in this case linear regression.
```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)

shape_lm <- lm(Diameter ~ RimType.general + RimAngle, 
                data = shape)
tab_model(shape_lm)

library(visreg)
visreg(shape_lm, "RimType.general", gg = TRUE) 
visreg(shape_lm, "RimAngle", gg = TRUE)
```

### Factor Analysis of Mixed Data
A good way of looking at all three variables together is to perform a Factor Analysis of Mixed Data (FAMD).
For resources, see: http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/115-famd-factor-analysis-of-mixed-data-in-r-essentials/

```{r}
library("FactoMineR")
library("factoextra")

shape.famd <- FAMD(shape, graph = FALSE)
```
The result is a list of values, which we can successively explore.
#### Eigenvalues
```{r}
eig.val <- get_eigenvalue(shape.famd)
head(eig.val)
fviz_screeplot(shape.famd)
```

The result shows that there is no high explanatory value for any one dimension. 

#### Variables
We will look at the variables now, and their correlation to the dimensions. We also look at their contribution of the variables to dimensions 1 and 2
```{r}
var <- get_famd_var(shape.famd)

## Coordinates of variables
head(var$coord)
## Cos2: quality of representation on the factor map
head(var$cos2)
## Contributions to the  dimensions
head(var$contrib)

## Plot of variables
fviz_famd_var(shape.famd, repel = TRUE)
## Contribution to the first dimension
fviz_contrib(shape.famd, "var", axes = 1)
## Contribution to the second dimension
fviz_contrib(shape.famd, "var", axes = 2)


```

We can see that RimType.general is the main contributor to both dimensions, whereas Diameter has more of an influence on Dim1 and none at all on Dim2. This reflects our previous observation that RimAngle and Diameter are not correlated with any noticeable strength.

#### Qualitative variables
Let's look at the two qualitative variables on their own, and see how their values behave.
```{r}
quali.var <- get_famd_var(shape.famd, "quali.var")
quali.var 

fviz_famd_var(shape.famd, "quali.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
             )
```

#### Individuals
Get results for the individual observations and thus approach groups.

```{r}
ind <- get_famd_ind(shape.famd)
ind
```

Now we will plot them, and color the groups according to the Rim Type
```{r}
fviz_mfa_ind(shape.famd, 
             habillage = "RimType.general", # color by groups 
             #palette = c("red", "green", "#FC4E07"),
             #addEllipses = TRUE, 
             repel = TRUE # Avoid text overlapping
             ) 
```

It is clearly visible now that, with a few exceptions, the general rim types are strongly related to Diameter and Angle of the rim. This is an important result. 
It means that, for this dataset at least, we can use the rim type as indicative of shape (on a statistical level). Interpretively, it means that vessels of different sizes also have different rims.

## Multiple Factor Analysis (MFA)
This is a way of looking at a large number of variables together, with the possibility of mixing categorical and numeric variables. This is not easy to do for decor, but we can leave it out and try a dataset without it.

```{r}
library(tidyr)
df_mfa <- select(rims, Site, Unit, Context, MaxTh, MinTh, Diameter, RimAngle, RimType, Burnish, Slip) 
#set Burnish and Slip to True or False
levels(df_mfa$Burnish) <- c(levels(df_mfa$Burnish), FALSE, TRUE)
df_mfa$Burnish[!is.na(df_mfa$Burnish)] = "TRUE"
df_mfa$Burnish <- replace_na(df_mfa$Burnish, FALSE)
as.logical(df_mfa$Burnish)

levels(df_mfa$Slip) <- c(levels(df_mfa$Slip), FALSE, TRUE)
df_mfa$Slip[!is.na(df_mfa$Slip)] = "TRUE"
df_mfa$Slip <- replace_na(df_mfa$Slip, FALSE)
as.logical(df_mfa$Slip)
```

Now we prepare for the factor analysis. The variables will be grouped here, according to what they represent.
Thus we group the first three variables (site, Unit, Context) into one group called location.

Next, we will group maximum and minimum thickness with diameter. We have already seen that they are correlated. We will specify that this group should be standardised (type s). We will leave the rim angle on its own, the rim type also, as well as burnish and slip.

NB: In future, I will see if there is a way to group variables of different kinds. Really, I would like to put max Thickness in with the Rim Type, and Rim Angle in with MIn THickness and Diameter, since this appears to make the most sense conceptually.

```{r}
library(FactoMineR)
res.mfa <- MFA(df_mfa, 
               group = c(3, 3,  1, 1, 1, 1), 
               type = c("n", "s", "n", "n", "n", "n"),
               name.group = c("location","dimensions", "shape",
                              "rim type", "burnish","slip"),
               num.group.sup = c(1),
               graph = FALSE)
print(res.mfa)
```
Check the eigenvalues to see whether this explains any variance:
```{r}
library("factoextra")
eig.val <- get_eigenvalue(res.mfa)
head(eig.val)
fviz_screeplot(res.mfa)
```
The answer to this is no. All the dimensions together explain only 10% of the variance in the dataset. This can be abandoned.

